Four Models for Severe Weather Deaths

### 2017-05-10 paper.Rnw
Added ZIP Code

\begin{thebibliography}{2}
  \bibitem{dataset} NOAA's Severe Weather Data Inventory, 
    \url{https://www1.ncdc.noaa.gov/pub/data/swdi/stormevents/csvfiles/}. 
    Accessed April 2017.
  \bibitem{ZIP} Bayarri, M., Berger, J., Datta, G. (2008). Objective testing of Poisson versus inflated Poisson models. IMS 	Collections, 3, 105-121. 
\end{thebibliography}
\clearpage
\appendix
\section{Code}
\label{a:code}
This appendix includes the code used to implement the models.
\subsection{Zero Inflated Negative Binomial code}
\label{a:cZINB}
<<zinbsampler-source, echo=TRUE, ref.label='zinbsampler'>>=
@

>>>>

\begin{thebibliography}{2}
  \bibitem{dataset} NOAA's Severe Weather Data Inventory, 
    \url{https://www1.ncdc.noaa.gov/pub/data/swdi/stormevents/csvfiles/}. 
    Accessed April 2017.
  \bibitem{ZIP} Bayarri, M., Berger, J., Datta, G. (2008). Objective testing of Poisson versus inflated Poisson models. IMS 	Collections, 3, 105-121. 
\end{thebibliography}
\clearpage
\appendix
\section{Code}
\label{a:code}
This appendix includes the code used to implement the models.
\subsection{Zero Inflated Negative Binomial code}
\label{a:cZINB}
<<zinbsampler-source, echo=TRUE, ref.label='zinbsampler'>>=
@
\subsection{Zero Inflated Poisson code}
<<eval = FALSE>>=
library(dplyr)
tornado <- data %>% filter(EVENT_TYPE == "Tornado") %>% select(DEATHS_DIRECT)
flood <- data %>% filter(EVENT_TYPE == "Flash Flood") %>% select(DEATHS_DIRECT)

# Precompute values to use in posterior function
zeros <- tornado %>% filter(DEATHS_DIRECT == 0) %>% select(DEATHS_DIRECT)
ones <- tornado %>% filter(DEATHS_DIRECT != 0) %>% select(DEATHS_DIRECT)
n <- nrow(ones)
t0_exp <- (zeros)/factorial(zeros)
l_t1_exp <- ones-1/2

# Define Posterior Functions
log_lam <- function(y, p, lambda){
  t0 <- sum(log(p + (1-p)*exp(-1*lambda)*(lambda+2e-10)^t0_exp))
  t1 <- sum(log(exp(-1*lambda)*lambda^l_t1_exp))
  return(t0+t1)
}

log_p <- function(y, p, lambda){
  t0 <- sum(log(p + (1-p)*exp(-1*lambda)*(lambda+2e-10)^t0_exp))
  t1_vec <- rep(1-p, n)
  t1 <- sum(log(t1_vec))
  return(t0+t1)
}

### Sampler for Tornado
set.seed(2011)

l_last <- 1
p_last <- .98

B <- 20000
lambdas <- rep(0, B)
ps <- rep(0, B)

l_accept <- rep(FALSE, B)
p_accept <- rep(FALSE, B)

for(t in 1:B){
  lambda <- rgamma(1, 3, 2)
  p <- rbeta(1, 1940, 60)
  ro_lam <- log_lam(tornado, p_last, lambda) - log(dgamma(lambda, 3, 2)) -
    log_lam(tornado, p_last, l_last) + log(dgamma(l_last, 3, 2))
  ro_p <- log_p(tornado, p, l_last) - log(dbeta(p, 1940, 60)) -
    log_p(tornado, p_last, l_last) + log(dbeta(p_last, 1940, 60))
  U_lambda <- runif(1)
  U_p <- runif(1)
  if(log(U_lambda) < min(0, ro_lam)){
    print(TRUE)
    lambdas[t] <- lambda
    lambda_last <- lambda
    l_accept[t] <- TRUE
  } else{
    print(FALSE)
    lambdas[t] <- lambda_last
  }

  if(log(U_p) < min(0, ro_p)){
    print(TRUE)
    ps[t] <- p
    p_last <- p 
    p_accept[t] <- TRUE
  } else{
    print(FALSE)
    ps[t] <- p_last
  }
}

# Do the same for flash floods

# Update values for likelihood function
zeros <- flood %>% filter(DEATHS_DIRECT == 0) %>% select(DEATHS_DIRECT)
ones <- flood %>% filter(DEATHS_DIRECT != 0) %>% select(DEATHS_DIRECT)
n <- nrow(ones)
t0_exp <- (zeros)/factorial(zeros)
l_t1_exp <- ones-1/2

# Sampler:
set.seed(2012)
l_last <- 1
p_last <- .98
B <- 20000
lambdas <- rep(0, B)
ps <- rep(0, B)
l_accept <- rep(FALSE, B)
p_accept <- rep(FALSE, B)
for(t in 1:B){
  print(paste0("t = ", as.character(t)))
  lambda <- rgamma(1, 1, 2)
  print(paste0("lambda = ", as.character(lambda)))
  p <- rbeta(1, 2945, 55)
  print(paste0("p = ", as.character(p)))
  ro_lam <- log_lam(flood, p_last, lambda) - log(dgamma(lambda, 1, 2)) -
    log_lam(flood, p_last, l_last) + log(dgamma(l_last, 1, 2))
  ro_p <- log_p(flood, p, l_last) - log(dbeta(p, 2945, 55)) -
    log_p(flood, p_last, l_last) + log(dbeta(p_last, 2945, 55))
  U_lambda <- runif(1)
  U_p <- runif(1)
  if(log(U_lambda) < min(0, ro_lam)){
    print(TRUE)
    lambdas[t] <- lambda
    lambda_last <- lambda
    l_accept[t] <- TRUE
  } else{
    print(FALSE)
    lambdas[t] <- lambda_last
  }

  if(log(U_p) < min(0, ro_p)){
    print(TRUE)
    ps[t] <- p
    p_last <- p 
    p_accept[t] <- TRUE
  } else{
    print(FALSE)
    ps[t] <- p_last
  }
}
@
